%% ============================================================
%% lss/adapter-contract.tex
%% ============================================================
%%
%% Canonical LSS adapter contract — what every LSS-X.go file must
%% guarantee.
%% ============================================================

\subsection*{LSS adapter contract}

\begin{definition}[LSS adapter contract]
\label{def:lss-adapter-contract}
An LSS adapter (e.g.\ $\mathsf{lss\_pulsar.go}$, $\mathsf{lss\_lens.go}$)
MUST satisfy:
\begin{enumerate}[nosep]
  \item \textbf{Lifecycle delegation.} The adapter does not implement
        share-secret arithmetic; it calls into the kernel
        ($\mathsf{Pulsar.Reshare}$, $\mathsf{Lens.Reshare}$, etc.) and
        wraps the output.
  \item \textbf{Generation increment.} On every successful
        $\mathsf{DynamicReshare}$, $g \mapsto g+1$ exactly. On
        $\mathsf{Rollback}$, $g \mapsto g_{\text{current}}+1$ with
        $\mathsf{rb} = g_{\text{current}}$.
  \item \textbf{KeyEraID preservation.} $\eta$ is read-only across
        $\mathsf{DynamicReshare}$. $\eta$ increments only on a separate
        $\mathsf{Reanchor}$ entry point.
  \item \textbf{GroupKey identity.} The kernel's $\mathsf{GroupKey}$
        pointer flows through unmodified.
  \item \textbf{Failure-mode containment.} On any verification failure
        (commit mismatch, activation rejection), return an error and
        retain the previous state. Do not silently fall back.
  \item \textbf{No ECDSA-specific assumptions.} Adapters for non-ECDSA
        kernels (Pulsar, Lens) MUST NOT inherit auxiliary secrets
        $w$, $q$ from the LSS-CMP path; the underlying kernel handles
        nonce blinding internally.
\end{enumerate}
\end{definition}

\begin{remark}[Adapter test contract]
The 10 acceptance tests in $\mathsf{lss\_pulsar\_test.go}$ and
$\mathsf{lss\_lens\_test.go}$ — GroupKey preservation, KeyEraID
preservation, Generation $+1$, RollbackFrom $= 0$ on forward, $t_{\text{old}}
\neq t_{\text{new}}$, disjoint set rotation, valid signature under
unchanged GroupKey, pairwise material regenerated, Rollback restores
prior state, malformed input rejected — collectively enforce
Definition~\ref{def:lss-adapter-contract} for the corresponding
adapter.
\end{remark}
