%% ============================================================
%% fhe/nebula-binding.tex
%% ============================================================
%%
%% Canonical Lux FHE result-root binding into Nebula/Quasar
%% transcripts. Papers cite this; do not redefine.
%% ============================================================

\subsection*{Lux FHE result roots bind into Nebula/Quasar transcripts}

\begin{theorem}[Nebula-FHE binding soundness]
\label{thm:fhe-nebula-binding}
Let $\mathsf{nebula\_fhe\_root}$ denote the canonical FHE evaluation
commitment (LP-167 \S\,Binding into Quasar finality):
\[
  \mathsf{nebula\_fhe\_root}
  \;=\;
  H\!\bigl(
    \mathsf{bootstrap\_key\_root}
    \;\|\;
    \mathsf{key\_switch\_key\_root}
    \;\|\;
    \mathsf{public\_param\_root}
    \;\|\;
    \mathsf{active\_circuit\_dag\_root}
    \;\|\;
    \mathsf{finalized\_evaluations\_root} \bigr)
\]
where $H$ is the Lux FHE hash suite (SHA-256 on rc1; SHA-3-256 on
production builds; suite identifier carried in the round descriptor per
LP-105 hash-suite invariance). Let $\mathsf{QuasarRoundDescriptor}$ bind
$\mathsf{nebula\_fhe\_root}$ alongside $\mathsf{NebulaRoot}$, $\mathsf{epoch}$,
$\mathsf{KeyEraID}$, $\mathsf{Generation}$, and $\mathsf{HashSuiteID}$
(LP-073, LP-105). Then for any PPT adversary $\mathcal{A}$ that produces a
\texttt{HorizonCertificate} accepting the descriptor, the probability that
$\mathcal{A}$ binds an FHE result not produced by an honest evaluation under
the keys committed by $\mathsf{bootstrap\_key\_root}$ and
$\mathsf{key\_switch\_key\_root}$ is bounded by
\[
  \mathrm{Adv}^{\mathsf{Nebula\text{-}FHE\text{-}bind}}_{\mathcal{A}}
  \;\leq\;
  \mathrm{Adv}^{\mathsf{coll}}_{H}(\mathcal{B}_1)
  \;+\;
  \mathrm{Adv}^{\mathsf{SUF\text{-}CMA}}_{\mathsf{Pulsar}}(\mathcal{B}_2)
  \;+\;
  \mathrm{negl}(\lambda)
\]
where $\mathcal{B}_1$ reduces to the collision resistance of $H$ and
$\mathcal{B}_2$ reduces to Pulsar SUF-CMA
(\texttt{proofs/pulsar/unforgeability.tex},
Theorem~\ref{thm:pulsar-suf-cma}).
\end{theorem}

\begin{proof}[Proof sketch]
A successful binding break either (i) collides $H$ on two distinct input
tuples to the same $\mathsf{nebula\_fhe\_root}$ --- reducing to the
collision-resistance assumption on the hash suite --- or (ii) forges a
Pulsar Pulse over the QuasarRoundDescriptor that contains the dishonest
$\mathsf{nebula\_fhe\_root}$, reducing to Pulsar SUF-CMA. The remaining
contribution is the negligible chance that the validator-set hash, the
KeyEra binding, or the Nebula transcript binding admit a parallel break;
each reduces by Prism's transcript-equality check (LP-105 \S Prism).
\end{proof}

\begin{remark}[F-Chain alias]
\label{rem:fchain-alias}
In F-Chain deployments (LP-013) the field $\mathsf{nebula\_fhe\_root}$
and $\mathsf{fchain\_fhe\_root}$ are the same 32-byte commitment; LP-013
names it from inside the F-Chain frame, LP-167 names it from the dual-
runtime perspective. The binding theorem applies unchanged; the
F-Chain cert lanes (\texttt{FChainTFHE}, \texttt{FChainBootstrap}) verify
against the same root.
\end{remark}

\begin{remark}[Replay protection]
\label{rem:fhe-replay-protection}
Cross-epoch replay of an FHE result is rejected because every contributing
sub-root ($\mathsf{bootstrap\_key\_root}$, $\mathsf{key\_switch\_key\_root}$,
$\mathsf{public\_param\_root}$, $\mathsf{active\_circuit\_dag\_root}$) is
keyed by $\mathsf{KeyEraID} + \mathsf{Generation}$, and the
$\mathsf{HashSuiteID}$ is bound into the transcript per LP-105
hash-suite separation. A descriptor referencing a stale
$\mathsf{nebula\_fhe\_root}$ fails Prism's transcript-equality check.
\end{remark}
